---
title: "ETC5523 Presentation"
subtitle: "<br/>Analysis of Melbourne City Council Climate Sensor Data"
author: 
 - "Vinny VU"
 - "Liu"
date: " `r Sys.Date()`"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
```

```{r, echo =FALSE, include = FALSE}
library(tidyverse)
library(lubridate)
library(leaflet)
library(visdat)
library(kableExtra)
library(tsibble)

readings <- read.csv("Microclimate_Sensor_Readings.csv")
locations <- read.csv("Microclimate_Sensor_Locations.csv")
```

```{r, echo = FALSE}
clean <- readings %>%
  select(site_id,sensor_id,local_time,type,value,units) %>%
  mutate(local_time = as_datetime(local_time)) %>%
  filter(type %in% c("TPH.RH")) %>%
  mutate(time = paste0(substr(local_time, 11,20)),
         date = date(local_time)) %>%
  mutate(time = hms(time)) %>%
  mutate(hour = hour(time))
  
```

# Background 

- Conducting an analysis of Melbourne City Council climate sensor data
- Two data sets: sensor locations, sensor readings 
- Data measures the climate of Melbourne sensors including humidity, temperature, and air quality
- Data can be found here: [sensor readings](https://data.melbourne.vic.gov.au/Environment/Microclimate-Sensor-Readings/u4vh-84j8?src=featured_banner), [Sensor locations](https://data.melbourne.vic.gov.au/Environment/Microclimate-Sensor-Locations/irqv-hjr4)
- **How does relative humidity in Melbourne change throughout the day and year?**
- **When is it most uncomfortable to walk around the city (humidity below 30% or above 60%)? **
- Relative humidity: the amount of water vapor present in air expressed as a percentage of the amount needed for saturation at the same temperature [howstuffworks](https://science.howstuffworks.com/nature/climate-weather/atmospheric/question651.htm)

???

---
class: inverse, center, middle

# Data Checking 

---

# Data Checking 

- Mean of arc1045 is significantly lower than the other sites
- Data for arc1048 is only collected until the 29th of August 

```{r, echo=FALSE, warning = FALSE, message = FALSE, fig.height=4, dev='svg'}

clean %>% 
  group_by(site_id) %>%
  summarise('Minimum Date'= min(date),
            'Maximum Date' = max(date),
            'Average Humidity' = round(mean(value, na.rm = TRUE),2)) %>%
  rename('Site ID' = site_id) %>%
  kable() %>%
  kable_styling()
```

---
# Data Checking

- There are missing values in the data set across 3 sites

```{r, echo = FALSE, warning = FALSE, message = FALSE}

miss <- clean %>%
  group_by(site_id,date) %>%
  summarise(value = mean(value))%>%
  as_tsibble(key = site_id, index = date)

gaps <- miss %>%
  count_gaps(.full = TRUE)
```

```{r, echo = FALSE, warning = FALSE, message = FALSE,fig.height=4, dev='svg'}
gaps %>%
  ggplot(aes(x = site_id, colour = site_id)) +
  geom_linerange(aes(ymin = .from, ymax = .to)) +
  geom_point(aes(y = .from)) +
  geom_point(aes(y = .to)) +
  labs(y = "", x = "", title = "Visualisation of Missing Values") +
  theme_minimal() +
  theme(legend.position = "none",
        text = element_text(size = 10),
        plot.title.position = "plot") +
  coord_flip()
```

---
# Locations 

- sensors: arc1045, arc1046, arc1048, arc1050 are all on Grattan Street
- sensor: arc1047 is on Pelham Street

```{r location, echo = FALSE, warning= FALSE, message= FALSE, dev='svg', out.width='100%', out.height= '70%'}
locations %>%
  leaflet() %>%
  addTiles() %>%
  addMarkers(~longitude,~latitude, popup = ~site_id)
```
---

class: inverse, center, middle

# Relative Humidity Across the Day

---
# Relative Humidity Across the Day by Site

- Site arc1045 shows significantly lower relative humidity levels across all hours of the day
- All other sensors show similar values with arc1047 showing higher peaks

```{r RH-day1, echo = FALSE, warning = FALSE, message = FALSE,fig.height=4, dev='svg'}
clean %>%
  group_by(site_id,sensor_id,type,hour) %>%
  summarise(value = mean(value, na.rm = TRUE)) %>%
  ggplot(aes(x = hour,
             y = value,
             colour = site_id)) +
  geom_line() +
  labs(x = "Hour",
       y = "Relative Humidity (%)",
       colour = "Site ID",
       title = "Average Relative Humidity by Hour by Site") +
  theme_minimal() +
  theme(legend.position = "bottom",
        text = element_text(size = 10),
        plot.title.position = "plot")
```

---

# Relative Humidity Across the Day Average

- Relative humidity peaks in the early morning at around 5am and decreases throughout the day dipping at around 2pm
- Perhaps when the sun is out vapor is evaporated throughout the day

```{r RH-day2, echo = FALSE, warning = FALSE, message = FALSE,fig.height=4, dev='svg'}
clean %>%
  group_by(hour) %>%
  summarise(value = mean(value, na.rm = TRUE)) %>%
  ggplot(aes(x = hour,
             y = value)) +
  geom_line(color = "steelblue") +
  labs(x = "Hour",
       y = "Relative Humidity (%)",
       colour = "Site ID",
       title = "Average Relative Humidity by Hour") +
  theme_minimal() +
  theme(text = element_text(size = 10),
        plot.title.position = "plot")
```

---

class: inverse, center, middle

# Relative Humidity Across the Year
---

# Relative Humidity Across the Year by Site

- Again we see arc1045 shows a consistently lower humidity levels across the year 
- All other sites show similar levels 

```{r RH-year1, echo = FALSE, warning = FALSE, message = FALSE,fig.height=4, dev='svg'}
clean %>%
  group_by(site_id,date) %>%
  summarise(value = mean(value, na.rm = TRUE)) %>%
  ggplot(aes(x = date,
             y = value,
             colour = site_id)) +
  geom_point(alpha = 0.5, size = 0.5) +
  geom_smooth(method = "loess", se = FALSE) +
  labs(x = "",
       y = "Relative Humidity (%)",
       colour = "Site ID",
       title = "Average Relative Humidity Across the Year by Site") +
  theme_minimal() +
  theme(legend.position = "bottom",
        text = element_text(size = 10),
        plot.title.position = "plot")
```
---

# Relative Humidity Across the Year by Average

- Humidity is lowest in the summer months and builds peaking in mid June
- In the summer the heat of the sun dries the vapor more compared to colder months  

```{r RH-year2, echo = FALSE, warning = FALSE, message = FALSE,fig.height=4, dev='svg'}
clean %>%
  group_by(date) %>%
  summarise(value = mean(value, na.rm = TRUE)) %>%
  ggplot(aes(x = date,
             y = value)) +
  geom_point(alpha = 0.5, size = 0.5) +
  geom_smooth(method = "loess", se = FALSE, color = "steelblue", alpha = 0.5, size = 0.5) +
  labs(x = "",
       y = "Relative Humidity (%)",
       colour = "Site ID",
       title = "Average Relative Humidity Across the Year") +
  theme_minimal() +
  theme(text = element_text(size = 10),
        plot.title.position = "plot")
```

---

# Comparing Across Sensors 

```{r site, echo = FALSE, warning = FALSE, message = FALSE,fig.height=5, dev='svg'}

med_ctr <- function(x, na.rm = TRUE) {
  x - mean(x, na.rm = TRUE)
}

clean %>% group_by(site_id,date) %>%
  filter(site_id != "arc1045") %>%
  summarise(value = mean(value, na.rm = TRUE)) %>%
  group_by(date) %>%
  mutate_at("value", med_ctr, na.rm = TRUE) %>%
  ggplot(aes(x = date,
             y = value,
             fill = site_id)) +
  geom_col() +
  facet_wrap(~site_id, ncol = 1) +
  labs(x = "",
       y = "Difference from Mean Relative Humidity (%)",
       title = "Average Differnce Relative Humidity across Sites") +
  theme_minimal() +
  theme(text = element_text(size = 10),
        legend.position = "none")
```

---

class: inverse, center, middle

# How does humidity effect US?

---

##  When is it most uncomfortable to walk around the city (humidity below 30% or above 60%)?

Hot weather:  
- Higher levels of humidity can make us feel even warmer
- The body is cooled through evaporation and higher levels of humidity can stifle this 
- On warm days with high levels of humidity we can feel warmer and "sticky" as our sweat does not evaporate as efficiently
- Think about countries near the equator such as Singapore, or Northern Queensland. 

[WeatherNation](https://www.weathernationtv.com/news/dry-cold-vs-humid-cold/)
---

##  When is it most uncomfortable to walk around the city (humidity below 30% or above 60%)?

Cold weather:  
- By the same logic on colder days with higher humidity we should feel warmer
- However, it is the opposite: we feel colder on cold days with higher humidity 
- High humidity impairs the insulation ability of clothing 
- Moisture in fibers and wicking effect of clothing can accelerate the heat loss through clothing 
- Humidity creates a thin layer of moisture on fibers and result in greater transfer of heat through clothing
- Think about rain on a cold day

[WeatherNation](https://www.weathernationtv.com/news/dry-cold-vs-humid-cold/)

#### It would be most uncomfortable to walk around the city on days with high humidity 

---

# References 
```{r, echo=FALSE, warning = FALSE, message = FALSE, dev='svg'}
reftab <- tibble(References = c("Microclimate Sensor Readings, City of Melbourne, (https://data.melbourne.vic.gov.au/Environment/Microclimate-Sensor-Readings/u4vh-84j8?src=featured_banner)",
                                "Microclimate Sensor Locations, City of Melbourne, (https://data.melbourne.vic.gov.au/Environment/Microclimate-Sensor-Locations/irqv-hjr4)",
                                "Welcome to the {tidyverse} Wickham, H., Averick, M., Bryan, J., Chang, W., McGowan, L.D., François, R., Grolemund, G., Hayes, A., Henry, L., Hester, J., Kuhn, M., Pedersen, T.L., Miller, E., Bache, S.M., Müller, K., Ooms, J., Robinson, D., Seidel, D.P., Spinu, V., Takahashi, K., Vaughan, D., Wilke, C., Woo, K. and Yutani, H., 2019. Journal of Open Source Software, Vol 4(43), pp. 1686. DOI: 10.21105/joss.01686",
                                "Dates and Times Made Easy with {lubridate} Grolemund, G. and Wickham, H., 2011. Journal of Statistical Software, Vol 40(3), pp. 1--25.",
                                "kableExtra: Construct Complex Table with 'kable' and Pipe Syntax     Zhu, H., 2019.",
                                "leaflet: Create Interactive Web Maps with the JavaScript 'Leaflet' Library (https://cran.r-project.org/web/packages/kableExtra/index.html) Cheng, J., Karambelkar, B. and Xie, Y., 2019.",
                                "visdat: Visualising Whole Data Frames (https://cran.r-project.org/web/packages/visdat/vignettes/using_visdat.html) Tierney, N., 2017. JOSS, Vol 2(16), pp. 355. Journal of Open Source Software. DOI: 10.21105/joss.00355",
                                "A new tidy data structure to support exploration and modeling of temporal data (https://github.com/tidyverts/tsibble) Wang, E., Cook, D. and Hyndman, R.J., 2020. Journal of Computational and Graphical Statistics, Vol 29(3), pp. 466-478. Taylor & Francis. DOI: 10.1080/10618600.2019.1695624",
                                "xaringan: Presentation Ninja, Yihui Xie, 2020, (https://CRAN.R-project.org/package=xaringan)",
                                "Dry Cold vs Humid Cold, WeatherNation, (https://www.weathernationtv.com/news/dry-cold-vs-humid-cold/)",
                                "What is Relative Humidity and How Does it Afffect How I Feel Outside, howstuffworks, (https://science.howstuffworks.com/nature/climate-weather/atmospheric/question651.htm)"))
  
reftab %>%
  kable() %>%
  kable_styling() %>%
  scroll_box(width = "100%", height = "500px")
```


---

class: inverse, center, middle

# Thanks!
