---
title: "ETC5523 Presentation"
subtitle: "<br/>Analysis of Melbourne City Council Climate Sensor Data"
author: 
 - "Vinny VU"
 - "Liu"
date: " `r Sys.Date()`"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
```

```{r, echo =FALSE, include = FALSE}
library(tidyverse)
library(lubridate)
library(leaflet)
library(visdat)
library(kableExtra)
library(tsibble)

readings <- read.csv("Microclimate_Sensor_Readings.csv")
locations <- read.csv("Microclimate_Sensor_Locations.csv")
```

```{r, echo = FALSE}
clean <- readings %>%
  select(site_id,sensor_id,local_time,type,value,units) %>%
  mutate(local_time = as_datetime(local_time)) %>%
  filter(type %in% c("TPH.RH")) %>%
  mutate(time = paste0(substr(local_time, 11,20)),
         date = date(local_time)) %>%
  mutate(time = hms(time)) %>%
  mutate(hour = hour(time))
  
```

# Background 

- Seek to understand climate sensor data from Melbourne City Council
- Two data sets: sensor locations, sensor readings 
- **How does relative humidity in Melbourne change throughout the day and year?**
- **When is it most uncomfortable to walk around the city (humidity below 30% or above 60%)? **
- Relative humidity: the amount of water vapor present in air expressed as a percentage of the amount needed for saturation at the same temperature

???

---
class: inverse, center, middle

# Data Checking 

---

# Data Checking 

```{r, echo=FALSE, warning = FALSE, message = FALSE, fig.height=4, dev='svg'}

clean %>% 
  group_by(site_id) %>%
  summarise('Minimum Date'= min(date),
            'Maximum Date' = max(date),
            'Average Humidity' = mean(value, na.rm = TRUE)) %>%
  rename('Site ID' = site_id) %>%
  kable() %>%
  kable_styling()
```

---
# Data Checking

- Missing values in the data set

```{r, echo = FALSE, warning = FALSE, message = FALSE}

miss <- clean %>%
  group_by(site_id,date) %>%
  summarise(value = mean(value))%>%
  as_tsibble(key = site_id, index = date)

gaps <- miss %>%
  count_gaps(.full = TRUE)
```

```{r, echo = FALSE, warning = FALSE, message = FALSE,fig.height=4, dev='svg'}
gaps %>%
  ggplot(aes(x = site_id, colour = site_id)) +
  geom_linerange(aes(ymin = .from, ymax = .to)) +
  geom_point(aes(y = .from)) +
  geom_point(aes(y = .to)) +
  labs(y = "", x = "", title = "Visualisation of Missing Values") +
  theme_minimal() +
  theme(legend.position = "none",
        text = element_text(size = 10),
        plot.title.position = "plot") +
  coord_flip()
```

---
# Locations 

```{r location, echo = FALSE, warning= FALSE, message= FALSE, dev='svg', out.width='100%'}
locations %>%
  leaflet() %>%
  addTiles() %>%
  addMarkers(~longitude,~latitude, popup = ~site_id)
```
---

class: inverse, center, middle

# Relative Humidity Across the Day

---
# Relative Humidity Across the Day by Site

```{r RH-day1, echo = FALSE, warning = FALSE, message = FALSE,fig.height=4, dev='svg'}
clean %>%
  group_by(site_id,sensor_id,type,hour) %>%
  summarise(value = mean(value, na.rm = TRUE)) %>%
  ggplot(aes(x = hour,
             y = value,
             colour = site_id)) +
  geom_line() +
  labs(x = "Hour",
       y = "Relative Humidity (%)",
       colour = "Site ID",
       title = "Average Relative Humidity by Hour by Site") +
  theme_minimal() +
  theme(legend.position = "bottom",
        text = element_text(size = 10),
        plot.title.position = "plot")
```

---

# Relative Humidity Across the Day Average

```{r RH-day2, echo = FALSE, warning = FALSE, message = FALSE,fig.height=4, dev='svg'}
clean %>%
  group_by(hour) %>%
  summarise(value = mean(value, na.rm = TRUE)) %>%
  ggplot(aes(x = hour,
             y = value)) +
  geom_line(color = "steelblue") +
  labs(x = "Hour",
       y = "Relative Humidity (%)",
       colour = "Site ID",
       title = "Average Relative Humidity by Hour") +
  theme_minimal() +
  theme(text = element_text(size = 10),
        plot.title.position = "plot")
```

---

class: inverse, center, middle

# Relative Humidity Across the Year
---

# Relative Humidity Across the Year by Site

```{r RH-year1, echo = FALSE, warning = FALSE, message = FALSE,fig.height=4, dev='svg'}
clean %>%
  group_by(site_id,date) %>%
  summarise(value = mean(value, na.rm = TRUE)) %>%
  ggplot(aes(x = date,
             y = value,
             colour = site_id)) +
  geom_point(alpha = 0.5, size = 0.5) +
  geom_smooth(method = "loess", se = FALSE) +
  labs(x = "",
       y = "Relative Humidity (%)",
       colour = "Site ID",
       title = "Average Relative Humidity by Day by Site") +
  theme_minimal() +
  theme(legend.position = "bottom",
        text = element_text(size = 10),
        plot.title.position = "plot")
```
---

# Relative Humidity Across the Year by Average

```{r RH-year2, echo = FALSE, warning = FALSE, message = FALSE,fig.height=4, dev='svg'}
clean %>%
  group_by(date) %>%
  summarise(value = mean(value, na.rm = TRUE)) %>%
  ggplot(aes(x = date,
             y = value)) +
  geom_point(alpha = 0.5, size = 0.5) +
  geom_smooth(method = "loess", se = FALSE, color = "steelblue", alpha = 0.5, size = 0.5) +
  labs(x = "",
       y = "Relative Humidity (%)",
       colour = "Site ID",
       title = "Average Relative Humidity by Day") +
  theme_minimal() +
  theme(text = element_text(size = 10),
        plot.title.position = "plot")
```

---

# Comparing Across Sensors 

```{r site, echo = FALSE, warning = FALSE, message = FALSE,fig.height=5, dev='svg'}

med_ctr <- function(x, na.rm = TRUE) {
  x - mean(x, na.rm = TRUE)
}

clean %>% group_by(site_id,date) %>%
  filter(site_id != "arc1045") %>%
  summarise(value = mean(value, na.rm = TRUE)) %>%
  group_by(date) %>%
  mutate_at("value", med_ctr, na.rm = TRUE) %>%
  ggplot(aes(x = date,
             y = value,
             fill = site_id)) +
  geom_col() +
  facet_wrap(~site_id, ncol = 1) +
  labs(x = "",
       y = "Difference from Mean Relative Humidity (%)",
       title = "Average Relative Humidity by Day by Site") +
  theme_minimal() +
  theme(text = element_text(size = 10),
        legend.position = "none")
```

---

class: inverse, center, middle

# How does humidity effect US?

---

##  When is it most uncomfortable to walk around the city (humidity below 30% or above 60%)?

Hot weather:  
- Higher levels of humidity can make us feel even warmer
- The body is cooled through evaporation and higher levels of humidity can stifle this 
- On warm days with high levels of humidity we can feel warmer and "sticky" as our sweat does not evaporate as efficiently
- Think about countries near the equator such as Singapore, or Northern Queensland. 

---

##  When is it most uncomfortable to walk around the city (humidity below 30% or above 60%)?

Cold weather:  
- By the same logic on colder days with higher humidity we should feel warmer
- However, it is the opposite: we feel colder on cold days with higher humidity 
- High humidity impairs the insulation ability of clothing 
- moisture in fibers and wicking effect of clothing can accelerate the heat loss through clothing 
- Humidity creates a thin layer of moisture on fibers on the fibers and result in greater transfer of heat through clothing
- Think about rain on a cold day

#### It would be most uncomfortable to walk around the city on days with high humidity 

---

class: inverse, center, middle

# Thanks!
